---
title: 'HealthCare HomeWork #2'
output:
  html_document:
    df_print: paged
---

#QUESTION 1

#### Read Data 
```{r, message=FALSE, warning=FALSE}
#rload packages
library(tidyverse);library(readxl); library(hhi)
setwd("~/My_R_Stuff/2019/Healthcare/HW2/Final HW2")
#Set uft-8
Sys.setlocale("LC_ALL", "en_US.UTF8")

#load plan data 
contract <-  read_excel("Monthly_Report_By_Plan_2019_01.xls")

#load enrollment data
enrollment <- read_csv("CPSC_Enrollment_Info_2019_01.csv", 
    col_types = cols(`Plan ID` = col_character()) )
#load dental data
dental = read_delim("pbp_b16_dental.txt", "\t", escape_double = FALSE, trim_ws = TRUE)

#load major insurance orgs 
MajorInsuranceOrgs <- read_excel("MajorInsuranceOrgs.xlsx")

ourstates = c("PA", "NJ", "IN", "SC", "CT", "MS", "HI", "SD" )

#load drug use data EOC170 measured by HEDIS
#EOC170 - Use of Opioids at High Dosage (UOD)
EOC170 <- read_excel("HEDIS2018.xlsx", 
    sheet = "EOC170")
```

#### Clean Enrollment Data

```{r, message=FALSE, warning=FALSE}
#convert enrollment data int
enrollment$Enrollment <- as.numeric(enrollment$Enrollment)

#select contracts that start with H,R, E
enrollment <- enrollment %>%
  filter(str_detect(`Contract Number`, "^H")| str_detect(`Contract Number`, "^R")|str_detect(`Contract Number`, "^E"))%>%
  select(`Contract Number`, `Plan ID`, State,Enrollment)

#filter out NAs
index <- which(is.na(enrollment$Enrollment)); enrollment <- enrollment[-index, ]

# summarize enrollment
enrollment <- enrollment %>%group_by(`Contract Number`, `Plan ID`, State) %>% summarise(Enrollment = sum(Enrollment))
```

#### Clean Contracts Data

```{r, message=FALSE, warning=FALSE}
# clean contracts data
contract <- contract %>%filter(str_detect(`Contract Number`, "^H")| str_detect(`Contract Number`, "^R")| 
         str_detect(`Contract Number`, "^E"))%>%select(`Contract Number`, `Plan ID`, `Organization Marketing Name`) 
```

#### Join Major Insurance Organization Data with Contracts Data

```{r, message=FALSE, warning=FALSE}
#join contracts and majorinsurance
contracts_insurance <- left_join(contract, MajorInsuranceOrgs, by = "Organization Marketing Name")

#join contracts_insurance with enrollments data and filter to teams states
final <- left_join(enrollment, contracts_insurance, by = c("Contract Number", "Plan ID")) %>% filter(State %in% ourstates)
```

#### calculate marketshare 
```{r}

marketshare <- final %>% ungroup%>% select(-`Contract Number`, -`Plan ID`, -`Organization Marketing Name`)%>%
  group_by(State, MajorInsuranceOrgName)%>% summarize(Enrollment = sum(Enrollment))

marketshare <- marketshare %>% ungroup %>% group_by(State)%>% mutate( MS = Enrollment / sum(Enrollment) * 100)
```


```{r}
#Calculate HHI
CT_1 <- marketshare %>% ungroup %>% filter(State == "CT")%>%select(MajorInsuranceOrgName, MS); CT <- round(hhi(data.frame(CT_1), "MS"), 0)
HI_1 <- marketshare %>% ungroup %>%filter(State == "HI")%>%select(MajorInsuranceOrgName, MS); HI <- round(hhi(data.frame(HI_1), "MS"), 0)
IN_1 <- marketshare %>% ungroup %>%filter(State == "IN")%>%select(MajorInsuranceOrgName, MS); IN <- round(hhi(data.frame(IN_1), "MS"), 0)
MS_1 <- marketshare %>% ungroup %>%filter(State == "MS")%>%select(MajorInsuranceOrgName, MS); MS <- round(hhi(data.frame(MS_1), "MS"), 0)
NJ_1 <- marketshare %>% ungroup %>%filter(State == "NJ")%>%select(MajorInsuranceOrgName, MS); NJ <- round(hhi(data.frame(NJ_1), "MS"), 0)
PA_1 <- marketshare %>% ungroup %>%filter(State == "PA")%>%select(MajorInsuranceOrgName, MS); PA <- round(hhi(data.frame(PA_1), "MS"), 0)
SC_1 <- marketshare %>% ungroup %>%filter(State == "SC")%>%select(MajorInsuranceOrgName, MS); SC <- round(hhi(data.frame(SC_1), "MS"), 0)
SD_1 <- marketshare %>% ungroup %>%filter(State == "SD")%>%select(MajorInsuranceOrgName, MS); SD <- round(hhi(data.frame(SD_1), "MS"), 0)
```

#### Make Table with State, hhi, MS and top insurance company 
```{r}

# make a dataframe
hhi_df = data.frame(rbind(CT, HI, IN, MS, NJ, PA, SC, SD)) %>% rename( HHI = rbind.CT..HI..IN..MS..NJ..PA..SC..SD.)

#vector of insurance companies
insurance <- c("UnitedHealthcare", "HMSA Akamai Advantage", "UnitedHealthcare", "Humana", "Aetna Health Inc.",
               "UPMC for Life", "UnitedHealthcare", "Medica Insurance Company")

#vector of insurance companies market shares
MS <- c(43.6, 28.6, 33.98, 57.2, 45.6, 13.2, 43.04, 65.3)
hhi_df_1 <- cbind(hhi_df, insurance, MS)

hhi_df_1
```

#QUESTION 2
#### Top 5 companies of market share for each state
```{r}
MS_top5 <- marketshare %>% ungroup %>% group_by(State) %>% top_n(5, MS)
```
#### clean the dental data
```{r}
dental = dental %>% select(`Contract Number` = pbp_a_hnumber, `Plan ID` = pbp_a_plan_identifier, segment_id, pbp_b16a_bendesc_yn, pbp_b16b_bendesc_yn)

dental_seg = dental %>% ungroup %>% 
  group_by(`Contract Number`, `Plan ID`) %>% 
  top_n(-1, segment_id)

```
##### merge by Orgcompanies
```{r,warning=FALSE}
dental_service = left_join(final,dental_seg, by=c("Contract Number", "Plan ID"))

dental_P <- dental_service %>% filter(pbp_b16a_bendesc_yn == 1) %>% group_by(State, MajorInsuranceOrgName) %>% summarise(Enroll_P=sum(Enrollment, na.rm = TRUE)) 
dental_C <- dental_service %>% filter(pbp_b16b_bendesc_yn == 1) %>% group_by(State, MajorInsuranceOrgName) %>% summarise(Enroll_C=sum(Enrollment, na.rm = TRUE))
dental_total <- dental_service %>% group_by(State, MajorInsuranceOrgName) %>% summarise(Enroll_T=sum(Enrollment, na.rm = TRUE))
#combine P C number
dental_total = left_join(dental_total,dental_P, by=c("State","MajorInsuranceOrgName"))
dental_total = left_join(dental_total,dental_C, by=c("State","MajorInsuranceOrgName"))
#calculate percentage
dental_total = dental_total %>% mutate(P_percentage = Enroll_P/Enroll_T,C_percentage = Enroll_C/Enroll_T) %>% select(State,MajorInsuranceOrgName,P_percentage, C_percentage)
#merge with top 5 in each state
dental_top5 = left_join(MS_top5, dental_total, by=c("State","MajorInsuranceOrgName"))
```

#QUESTION3

####Report the rate for EVERY CONTRACT offered by the top-10 biggest market share insurance companies in each state.
```{r}
#get the state-company-contract level enrollment 
test<- final %>% ungroup%>% select(-`Organization Marketing Name`)%>%
  group_by(`Contract Number`, State, MajorInsuranceOrgName)%>% summarize(Enrollment = sum(Enrollment))

#transform `EOC170-0010`from chr into numeric and filter out the missing values
EOC170$`EOC170-0010`<-as.numeric(EOC170$`EOC170-0010`)
index <- which(is.na(EOC170$`EOC170-0010`)); EOC170clean <- EOC170[-index, ]

#join test data and UOD rate data 
test_rate <- left_join(test, EOC170clean,  by = "Contract Number") %>% filter(!is.na(`EOC170-0010`))


#find the top-10 biggest market share insurance companies in each state
MS_top<-marketshare%>%ungroup%>% group_by(State) %>%top_n(10, MS)
MS_top<-MS_top %>% select(-Enrollment)
length(unique(MS_top$MajorInsuranceOrgName))#77 top 10 companies in total


#report the rate for EVERY CONTRACT offered by the top-10 biggest market share insurance companies in each state.

contract_top<-left_join(MS_top,test_rate,by = c("State", "MajorInsuranceOrgName")) %>% group_by(State,MajorInsuranceOrgName)%>% filter(!is.na(`EOC170-0010`))#filter the observations of the top 10 companies from each state

```




```{r}
# calculate the weighted average of the UOD rate based on the enrollment number of the company for the top-10 in each state can you
#for each contract, caculate ratebyenrollment = `EOC170-0010` * Enrollment
test_rate1 <- contract_top %>% mutate(ratebyenrollment = `EOC170-0010` * Enrollment)
#state-company level,sum the ratebyenrollment
test_rate2 <- test_rate1 %>% group_by(State, MajorInsuranceOrgName) %>% 
  mutate(totalratebyenrollment = sum(ratebyenrollment),
         companyUOD=ratebyenrollment/totalratebyenrollment) 
test_rate_rank<-test_rate2%>% group_by(State, MajorInsuranceOrgName)%>%arrange(desc(companyUOD), .by_group = TRUE)
```



